Different forms of uncertainty differently affect the evolution of social learning

### 2022-07-22 ms.tex
few paragraphs in to analysis revision

To address our research questions we developed an evolutionary agent-based model of
social learning evolution.  Evolutionary agent-based models simulate changes in
prevalence of a phenotype over several generations---in this model the phenotype, or
``trait'', is whether or not an agent engages in social learning.  In the model, a
population of $N$ individuals each must decide which of $B$ behaviors to perform at
each time step within a generation consisting of $L$ time steps. At the end of each
generation, agents are selected with replacement, biased by accumulated payoffs
within the generation, from the population to reproduce another $N$ agents. Agents
who inherit the social learning trait then learn about behavioral payoffs from the
previous generation, while asocial learners begin life with no \emph{a priori}
knowledge of the world.  Each behavior is a ``bandit'', a common modeling and
experimental approach for representing behaviors with Bernoulli-distributed
payoffs~\cite{SuttonBartoBook,McElreath2005,Steyvers2009,Rendell2010,Schulz2019}.
All bandits thus yield a payoff of 1 or 0 with a certain probability.  One of the
bandits/behaviors in the model is more likely to pay off than all the rest.  Agents
optimize their net payoffs over their lifespan by quickly figuring out which
behavior is optimal and performing it as often as possible within their lifespan.

We operationalized the four ``principal components'' of uncertainty that affect the
ability of agents to find and choose the optimal behavior. Recall these varieties of
uncertainty are \emph{environmental variability}; \emph{payoff ambiguity};
\emph{selection set size}; and the \emph{effective lifespan}. Environmental
variability is the probability that the behavior yielding the optimal payoff changes
from one generation to the next. We do not consider cases where the optimal behavior
varies within a generation. This is empirically justified since cultural evolution
occurs when meta-cognitive strategies and information are transferred via
learning and social influence, instead of by physical reproduction and death
as in genetic evolution.  The selection set size is the total number of possible
behaviors agents choose from, assumed to be constant within a simulation.  We
operationalized payoff ambiguity by manipulating the payoff difference between optimal
and non-optimal behaviors. The effective life span is the
number of time steps in which agents perform behaviors within a generation. 
Our model is not spatial, so agents have free, instant
access to all behaviors, and there is no spatial uncertainty, 
as might be the case for humans in a computer-based lab
experiment~\cite{McElreath2005,Morgan2012}.  Spatial dimensions could also be
considered irrelevant if costs associated with accessing a behavior were equal,
e.g., if it was assumed to be equally easy for bumblebees to fly to alternative
flower ``species''~\cite{Baracchi2018} or for great tits to fly to different food
sources~\cite{Aplin2017}.  Each simulated environment provides a distinct challenge
to the agents, which will lead agents to differently evolve to become social
learners or not, depending on which strategy optimizes net payoffs.

@@

In \emph{evolutionary} agent-based models like ours, agent life cycles are
programmed to cause agents to reproduce and die off, where reproducers pass on
heritable traits to their offspring. 
Natural selection occurs in the model when agents with higher
payoffs preferentially reproduce. This is a model of cultural evolution, so
``reproduction'' means teaching the meta-cognitive strategy of whether one should learn
socially at all; ``death'', in this interpretation, means
loss of the potential to socially influence others. 
This sets up an inter-generational structure
that we also use as a scaffold for social learning: we assume social learning occurs
\emph{obliquely}, meaning learning happens when a new
generation of agents each select a teacher from the previous generation, possibly
including their parents.  
Evolutionary agent-based models of cultural evolution are powerful, 
but their components need to be carefully specified and justified in relation to
empirical facts.

Evolution, then, acts as an optimization routine to identify which heritable trait
values are most rewarded under in different environmental uncertainty contexts,
defined by different uncertainty parameter settings
(Table~\ref{tab:uncertaintyParameters}).  The main outcome variable we analyze in
this model is therefore the value of agents' heritable trait, namely whether or not
an agent, $i$, learns socially, denoted $s_i$. $s_i$ is a binary variable: it is 1
if the agent is a social learner, and 0 otherwise. We will aggregate $s_i$ over all
agents and simulation trials to understand the affects of different uncertainty
parameter settings. We will support conclusions we make from observing patterns in
the outcomes with additional measures of the average time it takes for the
population to either all have $s_i = 1$ or all have $s_i = 0$, and the expected
payoffs in different contexts among populations that are initialized to all have
$s_i = 1$ or to all have $s_i = 0$. These outcome measures will be used to predict
the evolutionarily-favored strategy (to learn socially or not), the number of
generations it will take for a population to evolve the optimal strategy, and the
expected payoffs in populations performing the optimal strategy, across several specific
uncertainty contexts. To understand how different varieties of uncertainty interact
to affect the evolution of social learning, we systematically created an array of
simulated environments by systematically varying the uncertainty parameters.

@@

The model environment is dynamic: which behavior is optimal changes between
generations with probability equal to the \emph{environmental variability}, $u$.
The optimal behavior does not change within generations.
At the beginning of each generation, one bandit is optimal and yields expected
payoff $\pihigh$; the other $B-1$ behaviors yield a lower payoff, $\pilow$. 
The \emph{payoff ambiguity} here between optimal and non-optimal behavior
payoffs is defined by $\pihigh - \pilow$. We set $\pihigh = 0.9$ to be constant
across all simulations, so our analysis only varies payoff ambiguity by varying
$\pilow$. Specifically, payoff ambiguity increases with $\pilow$. 
By systematically varying these uncertainty parameters we change
the nature of the challenge faced by the agents. Each parameter individually 
can increase or decrease ``uncertainty,'' but it is not at all clear \emph{a priori}
how these uncertainties combine to affect the evolution of social learning.
These uncertainty parameters and the values we tested in our simulations
are summarized in Table~\ref{tab:uncertaintyParameters}.

@@

Agents are endowed with dynamic attributes to intelligently select behaviors and
calculate average payoffs for each behavior, and with a social learning trait value
that determines whether agents learn socially or not. All agents engage in
individual learning over their lifespan, whether or not they are social learners.
The difference between social learners and asocial learners is that social learners
have initially biased behavioral preferences based on information given to them by a
teacher of the previous generation. Asocial learners begin life with no behavioral
preferences. Agents and their learning procedures are specified through several
parameters introduced below.

@@

All agents perform individual, trial-and-error learning at each time step in
their lifespan.  This learning is guided by softmax search which
drives agents to more frequently exploit the most profitable behaviors when the
agent is more certain which are optimal, and to explore more frequently when the
agent is unsure.  To do softmax searching, agents track average payoffs acquired
from each behavior, which also requires agents to track the number of times they
have performed each behavior.  We denote the average payoffs obtained by agent $i$
for behavior $b$ as $\ledger$. The number of times $i$ performed $b$ is $c_{ib}$
(``$c$'' stands for count).

At each time step, agents perform behavior $b$ 
with softmax-weighted probability
\begin{equation}
  \Pr(\text{Agent $i$ performs behavior $b$}) = 
    \frac{e^{\beta \ledger}}{\sum_{b=1}^B e^{\beta \ledger}},
\end{equation}
\noindent
where $\beta$ is a parameter that adjusts how frequently agents perform 
behaviors with high expected payoffs (larger $\beta$) versus how frequently
agents explore alternative behaviors (smaller $\beta$)~\cite{McElreath2005}.

@@

At the beginning of each generation other than the first, social child agents select
one member of the previous generation to learn from if they are a social learner.
If agent $i$ is a social learner we write $s_i = 1$, and $s_i = 0$ otherwise.
A child agent $c$ is a social learner if its parent $p$ was a social learner,
i.e.\ $s_c \leftarrow s_p$, without mutation. 
Social learner children select a teacher from 
the previous generation by first selecting $N_T$ prospective teachers fully
at random, then by sampling from the $N_T$ with each prospective teacher weighted
by their net payoffs relative to the other $N_T - 1$.

@@

In this model, being a social learner is beneficial when it is more likely
on average to provide greater net payoffs than being an individual learner,
since evolution here acts as an optimization routine. 
Note that, unlike some models in the literature, it is not catastrophic for an 
individual to receive outdated information~\cite[e.g.]{Rogers1988}. Instead,
agents can explore alternatives and greedily perform a different observed optimal behavior
than an outdated optimal behavior they learned about socially. We rely
on this mechanism in the Analysis to explain the patterns we observe where
social learning evolves when $u > 0.5$, meaning that the
optimal behavior will switch between generations more often than not. If 
agents could not adjust to changing conditions like this, it seems like 
social learning would not evolve when $u > 0.5$.

@@

        $\beta$ & Softmax greediness; $\uparrow=$more exploitation, $\downarrow=$more
                    exploration 
               & 100, \textbf{10}, 1 \\

@@

\paragraph{Intragenerational dynamics}
% To guide behavior selection, agent $i$ counts how many times it
% performed each behavior $b$, denoted $c_{ib}$. Agents use this count to 
% update the mean payoff they have obtained from behavior $b$, denoted $\ledger$.
% $\ledger$ is initialized to 0 for all $b$ at model start for
% all agents (Figure~\ref{fig:schematic}A), and at 
% generation start for asocial learners. Social
% learning agents' $\ledger$ is initialized as that of its teacher, who is selected
% through performance-biased oblique learning; $c_ib$ is set to 1 for all social
% learners $i$ and behaviors $b$ (Figure~\ref{fig:schematic}C). Agent $i$'s
% accumulated payoffs from performing several behaviors over their lifetime of $L$
% time steps is denoted $\pi_{i}$. If an agent $i$ is a social learner, then we write
% $s_i = 1$, otherwise the agent is an asocial learner and $s_i = 0$.

@@

\paragraph{Intergenerational dynamics}
Between generations, agents first reproduce via asexual haploid reproduction; 
then new child agents learn from the
previous generation if they are social learners; and finally all agents from the
previous generation then die off (Figure~\ref{fig:schematic}C). 
$N$ reproducers are selected with replacement over $N$ independent draws, 
biased by performance:
\begin{equation}
  \Pr(\text{Agent $i$ is chosen to reproduce in one draw}) = \frac{\pi_i}{\sum_{i=1}^N \pi_i}.
\end{equation}
\noindent
A child inherits its parent's social learning trait $s_i$ without mutation.
A social learner child with $s_i = 1$ learns from a teacher from its parent's
generation, including possibly their parent (oblique learning). 
A child selects its teacher by first selecting $N_T$ prospective
teachers from the population completely at random, then selecting the one with
highest net payoffs $\pi_i$ as their teacher, with ties broken randomly. While it
is somewhat arbitrary to first subset $N_T$, this is a conservative choice that
represnts the fact that access to the entire population is not generally guaranteed.
Social learners adopt their chosen teacher's ledger of expected payoffs $\ledger$,
and set all $c_{ib} = 1$ so expected payoffs remain within $[0, 1]$.  Asocial
learners ($s_i = 0$) are initialized with all $c_{ib} = 0$ and $\ledger = 0$ 
(Figure~\ref{fig:schematic}C). The
newly initialized population then begins behavior selection and payoff accumulation
for another $L$ time steps.

After reproduction and die-off, the optimal behavior changes with probability
$u$. If a random draw dictates the behavior should change, the new
behavior is chosen from all other behaviors with the current optimal behavior
excluded.

@@

We developed a series of computational
analyses by systematically varying uncertainty parameters and observing the
frequency that model populations evolve to be social learners, the average payoffs
accrued by agents in each setting, and the time it takes for model populations to
reach fixation (agents all social or all asocial learners). Model dynamics are
illustrated in Figure~\ref{fig:schematic}.

@@

To analyze the effect of the four principal uncertainty factors, we systematically
varied their values and observed three key outcome variables aggregated across 1000
trial simulations, and across all agents when appropriate. We varied $u \in \{0.0,
0.1, \ldots, 1.0\}$; $\pilow \in \{0.1, 0.45, 0.8\}$; $B \in \{2, 4, 10\}$; and $L
\in \{1,2,4,8\}$ for $B=2$ and $L \in \{1,B/2,B,2B\}$ for $B=4,10$.  We observed
three outcome measures: (1) $\meansl$, the average value of $s_i$ over all agents
and trials; (2) $\meanpi / L$, the average net payoffs at simulation end across
agents and trials, normalized by lifespan; and (3) $\meanT / L$, 
or the mean number of time steps to fixation normalized by
lifespan. We analyze outcomes by plotting $\meansl$ on the y-axis and environmental
variation is on x-axis since we theoretically expect that $\meansl$ will 
decrease monotonically from 1 to 0 as $u$ increases. We expected and observed
that the other three uncertainty factors shift the value of $u$ at which 
$\meansl$ starts to decrease, and how rapidly $\meansl$ decreases. Note that the
hypothesis-testing concept of significance is meaningless here because we could
make any small outcome difference ``significant'' by running more simulation trials.
We instead study the patterns of outcome 
variables over systematically varied uncertainty factor values.

@@

We performed sensitivity analyses to ensure that our main analysis is reasonably
robust to a range of auxiliary parameters. Specifically in our sensitivity analyses
we first varied the population size, $N$, to explore whether smaller $N$ would 
show more drift due to finite population size effects, but also generally support
our main conclusions. Next, we varied the
cognitive model parameter $\beta \in \{1, 100\}$ to supplement the main results
that used $\beta = 10$. Recall that $\beta$ sets how frequently the softmax algorithm
exploited the best observed behavior versus explored the rest. We do not
expect agents with different $\beta$ to perform equally well 
individually, so certain $\beta$ values may themselves suppress social learning
since individual learning itself may be unreliable. Nonetheless our results
should be valid over at least some range of $\beta$. Finally, we varied the
number of prospective teachers $N_T \in \{2, 10, 20\}$ 
to supplement the main results that used $N_T = 5$. The number of prospective
teachers should not change the whether social learning is optimal or not, 
but it may make model results driftier since the benefit of social learning should
be more difficult to detect with fewer prosopective teachers.

@@

To understand how different forms of uncertainty interact to affect the evolution
of social learning, we measured simulated social learning prevalence across a 
suite of simulations over uncertainty parameters. We analyze simulation outcomes
by plotting the mean observed social learning frequency, $\meansl$, across all agents 
and 1000 simulation trials for each uncertainty parameter setting. 
We visualize and inspect plots with $\meansl$ on the y-axis and
environmental variability, $u$ on the x-axis. This is convenient because we are
most certain about how $\meansl$ depends on $u$: we expect $\meansl$ to decrease
as $u$ increases since environmental variability is well understood to cause
socially-transmitted information to become outdated. When the other forms of
uncertainty are varied, this causes changes to when and how quickly social learning
is suppressed as $u$ increases. To inspect these changes, we show nine plots
of $\meansl$ over $u$ in a $3\times3$ plot array (Figure~\ref{fig:mainResults}). 
Plot array columns correspond to different selection set sizes, $B$; rows
correspond to different non-optimal payoff values, $\pilow$ (recall optimal
payoff $\pihigh=0.9$ for all analyses); plot line and marker colors correspond
to different effective lifespan, $L$, values. 
We support our conclusions regarding the evolution of
social learning by analyzing the time to 
fixation and by comparing the average net payoffs in our simulations to the
expected payoffs for individuals in both populations of all social learners
and in populations of all individual learners. We thus use our analysis 
to understand changes to how $\meansl$ changes over $u$ for different uncertainty parameter settings.

Several patterns emerged in changes to social learning prevalence over
as the uncertainty parameters were varied (Figure~\ref{fig:mainResults}). 
First, we indeed observed that increased $u$ uniformly led to decreased
$\meansl$, across the other uncertainty parameter values. The other uncertainty
parameters changed at which value of $u$ that $\meansl$ started to decrease,
and how quickly $\meansl$ decreased. For example, 
when payoff ambiguity and selection set size were
both small ($\pilow = 0.1$ and $B=2$) social learning only evolved when $u$ is small
since individual learning is relatively efficient in this case
(Figure~\ref{fig:mainResults}, upper left). As $\pilow$ increased,
$\meansl$ decreased less rapdily over $u$, which was more pronounced for smaller $L$ (rows of
Figure~\ref{fig:mainResults}).

Increased selection set size $B$ led nearly uniformly to more prevalent social
learning with all other uncertainty parameters being equal ($B$ increases along
columns in Figure~\ref{fig:mainResults}. 
This is because possibly misleading 
social information is less consequential due
to (1) larger $B$ is a more difficult problem for trial-and-error learning, and
(2) a misidentified optimal behavior is weighted less relative to smaller
$B$, meaning agents are more likely to explore behaviors identified as non-optimal
when $B$ is larger (columns of Figure~\ref{fig:mainResults}).  \mt{Support this by calculating the difference between weights when B=2 and B=10 and a social learner learns outdated information}
Increased $B$ led to more pronounced flattening of social learning suppression when
$L$ was small, apparently again because of the limited data available to evolution,
making individual outcomes more path dependent.

Payoff ambiguity had a modest effect when $\pilow = 0.1$ increased to $\pilow = 0.45$,
causing social learning to be favored across more values of $u$, across all
selection set sizes and lifespans (first two rows of Figure~\ref{fig:mainResults}). 
This indicates that increased payoff ambiguity significantly increases the 
difficulty of discerning which behavior is optimal since agents evolve to 
risk social learning at greater values of $u$.
When $\pilow = 0.8$ the largest effect is smoothing and flattening of the $\meansl$ curves,
which indicates more drift (bottom row of Figure~\ref{fig:mainResults}). 
Initially this results in less frequent evolution
of social learning for the same values of $u$ and $B$. But after 
the $u$ value where $\meansl \approx 0.5$, social learning evolves more frequently than
for the same $u$, $B$, and $L$. When $\pilow \to \pihigh$, agents do better and better
whether they are social learners or not. We see that this leads to less penalty
for trying social information before possibly having to correct course due to
outdated social information.

>>>>

To understand how uncertainty affects social learning, 
we developed an agent-based model of
social learning evolution.  Evolutionary agent-based models simulate changes in
prevalence of a phenotype over several generations---in this model the phenotype
is whether or not an agent engages in social learning.  In the model, a
population of $N$ individuals each must decide which of $B$ behaviors to perform at
each time step within a generation consisting of $L$ time steps. At the end of each
generation, agents are selected with replacement to reproduce another $N$ agents, 
biased by net payoffs. Agents
who inherit the social learning trait then learn about payoffs from the
previous generation, while asocial learners begin life with no \emph{a priori}
knowledge of the world.  Each behavior is a ``bandit'', a common modeling and
experimental approach for representing behaviors with Bernoulli-distributed
payoffs~\cite{SuttonBartoBook,McElreath2005,Steyvers2009,Rendell2010,Schulz2019}.
All bandits thus yield a payoff of 1 or 0 with a certain probability.  One of the
bandits/behaviors in the model is more likely to pay off than all the rest.  Agents
optimize their net payoffs over their lifespan by quickly figuring out which
behavior is optimal and performing it as often as possible within their lifespan.

We operationalized four separable varieties of uncertainty that affect the
ability of agents to find and choose the optimal behavior: 
\emph{environmental variability}; \emph{payoff ambiguity};
\emph{selection set size}; and the \emph{effective lifespan}. Environmental
variability is the probability that the optimal behavior changes
between generations. We do not consider within-generation variability. 
Furthermore, our model is not spatial, so agents have free, instant
access to all behaviors, and there is no spatial uncertainty. 
These conditions reflect humans in a computer-based social learning evolution
experiment~\cite{McElreath2005,Morgan2012,Derex2016}. Spatial dimensions could also be
considered irrelevant if costs associated with accessing a behavior were equal,
e.g., if it was assumed to be equally costly for animals to
access alternative food sources~\cite{Aplin2017,Baracchi2018}.  
% This is empirically justified since cultural evolution
% occurs when meta-cognitive strategies and information are transferred via
% learning and social influence, instead of by physical reproduction and death
% as in genetic evolution.  
The selection set size is the total number of possible
behaviors agents choose from, assumed to be constant within a simulation.  
Payoff ambiguity is the payoff difference between optimal
and non-optimal behaviors. The effective life span is the
number of time steps in which agents perform behaviors within a generation. 
Each simulated environment provides a distinct challenge
to the agents, which will lead agents to differently evolve to become social
learners or not, depending on which strategy optimizes net payoffs.

@@

In \emph{evolutionary} agent-based models like ours, agent life cycles are
programmed to cause agents to reproduce and die off, where reproducers pass on
heritable traits to their offspring. 
Natural selection occurs in the model when agents with higher
payoffs preferentially reproduce. This is a model of cultural evolution, so
``reproduction'' means teaching the meta-cognitive strategy of whether one should learn
socially at all; ``death'', in this interpretation, means
loss of the potential to socially influence others. 
This sets up an inter-generational structure
that we also use as a scaffold for social learning: we assume social learning occurs
\emph{obliquely}, meaning learning happens when a new
generation of agents each select a teacher from the previous generation, possibly
including their parents.  
Evolution, then, acts as an optimization routine to identify which heritable trait
values are most rewarded under in different environmental uncertainty contexts,
defined by different uncertainty parameter settings
(Table~\ref{tab:uncertaintyParameters}).

Agents have one heritable trait in this model, whether or not
an agent, $i$, learns socially, denoted $s_i$. $s_i$ is binary: it is 1
if the agent is a social learner, and 0 otherwise. In our analysis, we 
aggregate $s_i$ over all agents and simulation trials to understand the affects of
different uncertainty parameter settings. 
Our computational analysis amounts to observing patterns in
the aggregated $s_i$, supported by measuring the number of generations 
to model convergence and expected payoffs. 
To understand how different varieties of uncertainty interact
to affect the evolution of social learning, we systematically created an array of
simulated environments by systematically varying the uncertainty parameters.

@@

The model environment is dynamic: which behavior is optimal changes between
generations with probability equal to the \emph{environmental variability}, $u$.
At the beginning of each generation, one bandit is optimal and yields expected
payoff $\pihigh$; the other $B-1$ behaviors yield a lower payoff, $\pilow$. 
The \emph{payoff ambiguity} is defined by $\pihigh - \pilow$. 
We set $\pihigh = 0.9$, constant across all simulations, so we vary
payoff ambiguity by varying $\pilow$, i.e., payoff ambiguity increases with
$\pilow$.

@@

Agents are endowed with dynamic attributes to intelligently select behaviors and
calculate average payoffs for each behavior. All agents engage in
individual learning over their lifespan, whether or not they are social learners.
Social learners begin life with
behavioral preferences based on information from their
teacher, an agent from the previous generation. 
Asocial learners begin life with no behavioral
preferences. 
% Agents and their learning procedures are specified through several
% parameters introduced below.

@@

All agents perform individual, trial-and-error learning at each time step in
their lifespan.  Learning is guided by softmax search. Softmax search
guides agents to more frequently exploit the most profitable behaviors when the
agent is more certain which is optimal, and to explore more frequently when the
agent is unsure.  To do softmax searching, agents track average payoffs acquired
from each behavior, which also requires agents to track the number of times they
have performed each behavior.  We denote the average payoffs obtained by agent $i$
for behavior $b$ as $\ledger$. We denote the number of times $i$ performed $b$
as $c_{ib}$ (``$c$'' stands for count). 
At each time step, agents perform behavior $b$ with softmax-weighted probability
\begin{equation}
  \Pr(\text{Agent $i$ performs behavior $b$}) = 
    \frac{e^{\beta \ledger}}{\sum_{b=1}^B e^{\beta \ledger}}.
\end{equation}
\noindent
$\beta$ adjusts how frequently agents perform 
behaviors with high expected payoffs (larger $\beta$) versus how frequently
agents explore alternative behaviors (smaller $\beta$)~\cite{McElreath2005}.

@@

At the beginning of each generation other than the first, social child agents select
one member of the previous generation to learn from, if the child inherited
$s_i = 1$. A child agent $c$ is a social learner if its parent $p$ was a social
learner, i.e.\ $s_c \leftarrow s_p$, without mutation.  Social learner children
select a teacher from the previous generation by first selecting $N_T$ prospective
teachers fully at random, then by sampling from the $N_T$ with each prospective
teacher weighted by their net payoffs relative to the other $N_T - 1$.

@@

In this model, being a social learner is beneficial when it is more likely on
average to provide greater net payoffs than being an individual learner.  It is not
catastrophic for an individual to receive outdated information, as is sometimes
assumed~\cite[e.g.]{Rogers1988}. Instead, agents explore alternatives and
greedily perform different behaviors that pay off better. 
This mechanism is critical for explaining the patterns we observe in the 
Analysis, where social learning evolves when $u > 0.5$.
In this case, the optimal behavior will switch between generations more often than
not. Without individual learning, social learning would not evolve when $u > 0.5$.

@@

        $\beta$ & Softmax greediness; $\uparrow=$more exploitation, $\downarrow=$more
                    exploration 
               & 1, \textbf{10}, 100 \\

@@

\paragraph{Within-generation dynamics}
% To guide behavior selection, agent $i$ counts how many times it
% performed each behavior $b$, denoted $c_{ib}$. Agents use this count to 
% update the mean payoff they have obtained from behavior $b$, denoted $\ledger$.
% $\ledger$ is initialized to 0 for all $b$ at model start for
% all agents (Figure~\ref{fig:schematic}A), and at 
% generation start for asocial learners. Social
% learning agents' $\ledger$ is initialized as that of its teacher, who is selected
% through performance-biased oblique learning; $c_ib$ is set to 1 for all social
% learners $i$ and behaviors $b$ (Figure~\ref{fig:schematic}C). Agent $i$'s
% accumulated payoffs from performing several behaviors over their lifetime of $L$
% time steps is denoted $\pi_{i}$. If an agent $i$ is a social learner, then we write
% $s_i = 1$, otherwise the agent is an asocial learner and $s_i = 0$.

@@

\paragraph{Intergenerational dynamics}
Between generations, agents first reproduce via asexual haploid reproduction; 
then new child agents learn from the
previous generation if they are social learners; and finally all agents from the
previous generation then die off (Figure~\ref{fig:schematic}C). 
$N$ reproducers are selected with replacement over $N$ independent draws, 
biased by performance:
\begin{equation}
  \Pr(\text{Agent $i$ is chosen to reproduce in one draw}) = \frac{\pi_i}{\sum_{i=1}^N \pi_i}.
\end{equation}
\noindent
A child inherits its parent's social learning trait $s_i$ without mutation.
A social learner child with $s_i = 1$ learns from a teacher from its parent's
generation, including possibly their parent (oblique learning). 
A child selects its teacher by first selecting $N_T$ prospective
teachers from the population completely at random, then selecting the one with
highest net payoffs $\pi_i$ as their teacher, with ties broken randomly. While it
is somewhat arbitrary to first subset $N_T$, this is a conservative choice that
represnts the fact that access to the entire population is not generally guaranteed.
Social learners adopt their chosen teacher's expected payoffs observations $\ledger$,
and set all $c_{ib} = 1$ so expected payoffs remain within $[0, 1]$.  Asocial
learners ($s_i = 0$) are initialized with $c_{ib} = 0$ and $\ledger = 0$ 
(Figure~\ref{fig:schematic}C). The
newly initialized population then repeats within-generation dynamics.

After reproduction and die-off, the optimal behavior changes with probability
$u$. The new behavior is chosen from all other behaviors, with the 
current optimal behavior excluded from selection.

@@

% We developed a series of computational
% analyses of the effect of uncertainty on social learning evolution by 
% systematically varying uncertainty parameters and observing how
% frequently model populations evolve to be social learners.

To analyze the effect of the four principal uncertainty factors, we systematically
varied uncertainty values and observed how frequently social learning evolved across
1000 trial simulations. We varied $u \in
\{0.0, 0.1, \ldots, 1.0\}$; $\pilow \in \{0.1, 0.45, 0.8\}$; $B \in \{2, 4, 10\}$;
and $L \in \{1,2,4,8\}$ for $B=2$ and $L \in \{1,B/2,B,2B\}$ for $B=4,10$.  We
observed three outcome measures, with $\meansl$ as the primary measure: 
the average value of $s_i$ over all agents and trials. To support our conclusions
based on $\meansl$, we also measured $\meanpi / L$, the average net 
payoffs at simulation end across agents and trials, normalized by lifespan; 
and $\meanT / L$, or the mean
number of time steps to fixation normalized by lifespan. We analyze outcomes by
plotting $\meansl$ on the y-axis and environmental variation is on x-axis since we
theoretically expect that $\meansl$ will decrease monotonically from 1 to 0 as $u$
increases. Note that the hypothesis-testing concept of significance is meaningless
here because we could make any small outcome difference ``significant'' by running
more simulation trials.  We instead study the patterns of outcome variables over
systematically varied uncertainty factor values.

@@

@@

We performed sensitivity analyses to ensure that our main analysis is reasonably
robust to a range of auxiliary parameters. Specifically,
we varied the population size, $N$, to explore whether smaller $N$ would 
show more drift due to finite population size effects, but also generally support
our main conclusions. Next, we varied the
softmax greediness parameter $\beta \in \{1, 100\}$ to supplement the main results
that used $\beta = 10$. We do not
expect agents with different $\beta$ to perform equally well 
individually, so certain $\beta$ values may themselves suppress social learning
by undermining individual learning. Nonetheless, our results
should be valid over some range of $\beta$. Finally, we varied the
number of prospective teachers $N_T \in \{2, 10, 20\}$ 
to supplement the main results that used $N_T = 5$. The number of prospective
teachers should not change whether social learning is optimal, 
but it may induce more drift since the benefit of social learning may
be more difficult to detect with fewer prosopective teachers.

@@

To understand how different forms of uncertainty interact to affect the evolution of
social learning, we measured how frequently social learning evolved across
uncertainty parameters. We analyze outcomes by plotting the mean observed social
learning frequency, $\meansl$, across all agents and 1000 simulation trials for each
uncertainty parameter setting.  We visualize and inspect plots with $\meansl$ on the
y-axis and environmental variability, $u$ on the x-axis. 
% This is convenient because we are most certain about how $\meansl$ depends on $u$:
% we expect $\meansl$ to decrease as $u$ increases since environmental variability
% is well understood to cause socially-transmitted information to become outdated. 
To inspect patterns in $\meansl$ over all uncertainty parameters, we analyze nine
plots of $\meansl$ over $u$ in a $3\times3$ plot array
(Figure~\ref{fig:mainResults}).  Plot array columns correspond to different
selection set sizes, $B$; rows correspond to different ambiguous payoffs via
non-optimal expected payoffs, $\pilow$ (recall $\pihigh=0.9$ for all analyses);
finally, plot line and marker colors correspond to different effective lifespans,
$L$.  
% We support our conclusions regarding the evolution of social learning by analyzing
% the time to fixation and by comparing the average net payoffs in our simulations
% to the expected payoffs for individuals in both populations of all social learners
% and in populations of all individual learners. We thus use our analysis to
% understand changes to how $\meansl$ changes over $u$ for different uncertainty
% parameter settings.

Several patterns emerged in changes to social learning prevalence over
the uncertainty parameters (Figure~\ref{fig:mainResults}). 
First, we indeed observed that greater $u$ uniformly decreased
$\meansl$, across all other uncertainty parameter values. The other uncertainty
parameters changed when $\meansl$ began to decrease in terms of $u$,
and how quickly $\meansl$ decreased from 1 to 0 over $u$. For example, 
when payoff ambiguity and selection set size were
both small ($\pilow = 0.1$ and $B=2$) social learning only evolved when $u$ was
small.
This is a relatively easy problem, and individual learning is more
efficient in this case (Figure~\ref{fig:mainResults}, upper left).
As $\pilow$ increased, $\meansl$ decreased less rapdily over $u$, 
which was more pronounced for smaller $L$ (see rows of Figure~\ref{fig:mainResults}).

Increased selection set size $B$ led nearly uniformly to more prevalent social
learning with all other uncertainty parameters being equal ($B$ increases along
columns in Figure~\ref{fig:mainResults}).  This is because misleading social
information is easy to overcome for two reasons: (1) larger $B$ presents a more
difficult problem for trial-and-error learning; and (2) a misidentified optimal
behavior is weighted less relative to smaller $B$, meaning agents are more likely to
explore behaviors identified as non-optimal when $B$ is larger (columns of
Figure~\ref{fig:mainResults}).  
To understand point (2), consider the case where a social learner child learns
that behavior 1 has an expected payoff of 0.8, and the rest have an expected
payoff of 0, but the optimal behavior changes between generations. Assume
the child tries behavior 1, but it pays off 0.
If $B=2$, then at within-generation $t=2$, $\ledger = (0.4, 0.0)$. The probability
the agent explores behavior 2 is now 0.02. However, if $B=10$ the agent has
$\ledger = (0.4, 0.0, \ldots, 0.0)$; the probability
the agent explores one of the nine alternative behaviors increases to 0.14.
If the agent again tries behavior 1, the probability of exploring alternatives
with $B=2$ is 0.21, but with $B=10$ the probability of exploring alternatives
greatly increases to 0.71.
% Larger $B$ caused more drift when
% $L$ was small, apparently due to the limited data available to evolution.

Payoff ambiguity had a modest effect when $\pilow = 0.1$ increased to $\pilow =
0.45$, causing social learning to be favored across more values of $u$, and across
all selection set sizes and lifespans (first two rows of
Figure~\ref{fig:mainResults}).  This indicates that increased payoff ambiguity
increases the difficulty of discerning which behavior is optimal since agents evolve
to risk social learning at greater values of $u$.  When $\pilow = 0.8$ the largest
effect is smoothing and flattening of the $\meansl$ curves, which indicates more
drift (bottom row of Figure~\ref{fig:mainResults}).  Initially this results in less
frequent evolution of social learning for the same values of $u$ and $B$. But after
the $u$ value where $\meansl \approx 0.5$, social learning evolves more frequently
than for the same $u$, $B$, and $L$. When $\pilow \to \pihigh$, agents do better and
better whether they are social learners or not. We see that this leads to less
penalty for trying social information before possibly having to correct course due
to outdated social information.
